# MPD summaries



## Single Model Output
```{r read_single_model, fig.width=6, fig.height=4, results = 'hide', message=FALSE, warning=FALSE}
file_name = system.file("extdata","SimpleTestModel", 
                        "estimate.log", package = "r4Casal2", mustWork = TRUE)
mpd = extract.mpd(file = file_name)
```

### Model Convergence 
There are a range of appproaches for checking your model has converged. The approaches we will be working through include checking the hessian is positive definite, checking parameters are not running to bounds and reestimate with random starting locations.

When estimating models in Casal2, it is recommended to have the following report included 
```{bash covar_report, eval = F}
@report covariance_matrix
type covariance_matrix
## or the Hessian
@report hessian_matrix
type hessian_matrix
```


When estimation is complete and you have read in the Casal2 output using `Casal2::extract.mpd()`. 

```{r check covariance, fig.width=6, fig.height=4, results = 'hide', message=FALSE, warning=FALSE}
# file name
mpd_file_name = system.file("extdata", "PosteriorPredictiveChecks","estimate.log", 
                            package = "r4Casal2", mustWork = TRUE)
# read in output
mpd = extract.mpd(file = mpd_file_name)
# is covariance symetric
isSymmetric(mpd$covar$covariance_matrix)
# is hessian invertable
is_matrix_invertable(mpd$hess$hessian_matrix)
# check high correlations
correlation_matrix = cov2cor(mpd$covar$covariance_matrix)
corr_df = get_high_correlations(correlation_matrix = correlation_matrix, max_correlation = 0.8, 
                      labels = names(mpd$estimate_value$values))
corr_df$correlation = round(corr_df$correlation, 3)
```

```{r correlated_params}
kable(x = corr_df[,c("correlation", "row_param", "col_param")], 
      caption = "Correlated Parameters")
```
You will want to try remove high correlations from the covariance to help estimation and MCMC simulations. We recommend you explore parameter transformations to remove high correlations or alternative parameterisations.


Once these are satisfied you will have more confidence in your standard errors, in addition to being able to run MCMC run mode.

Another useful convergence diagnostic is re-estimating Casal2 with difference starting locations. The function used for this is `?generate.starting.pars`. This will read a Casal2 config file that contains all the `@estimate` definitions are generate a bunch of random starting values in the format of useable for `-i` in Casal2. Below is some example R code of running Casal2 from R with randomly generated starting values.


```{r code_to_jitter,eval=F}
working_dir = "Directory to Casal output"
## generate starting values
start_pars = generate.starting.pars(path = working_dir, 
                                    estimation_csl2_file  = "estimation.csl2",
                                    par_file_name = "starting_pars.out")
## re-run Casal2
current_dir = getwd()
setwd(working_dir)
system2(command = "casal2", args = "-e -o multi_start_pars.par -i starting_pars.out",
        stdout = "multi_start.log",
        stderr = "multi_start.err", wait=TRUE)
system2(command = "casal2", args = "-r -i starting_pars.out",
        stdout = "multi_start_init.log",
        stderr = "multi_start_init.err", wait=TRUE)
setwd(current_dir)

## read in jitter_start run
multi_est = extract.mpd("multi_start.log", path = working_dir)
multi_run = extract.mpd("multi_start_init.log", path = working_dir)
## check if any didn't converge

## plot SSBS
ssb_df = get_derived_quanitites(multi_est)
ggplot(ssb_df, aes(x = years, y = values, col = par_set, linetype = par_set)) +
  geom_line(size = 1.5) +
  labs(x = "Years",y = "SSB", linetype = "Starting\nvalues", col = "Starting\nvalues")
ssb_df = get_derived_quanitites(multi_run)
ggplot(ssb_df, aes(x = years, y = values, col = par_set, linetype = par_set)) +
  geom_line(size = 1.5) +
  labs(x = "Years",y = "SSB", linetype = "Starting\nvalues", col = "Starting\nvalues")

## get aggregated objective functions
obj = aggregate_objective_report(model = multi_est)
head(obj)
```



### Data Weighting
Some pseudo r code to help with data weighting according to @francis2011data with multinomial data.
```{r data_weighting,eval=F}
working_dir = "Directory to Casal output"
reweight_folder = "Reweight"
## Don't always want to re-run this code
if(FALSE) {
  weights = run_automatic_reweighting(config_dir = working_dir,
                                      config_filename = "config.csl2",
                                      weighting_folder_name = reweight_folder,
                                      mpd_file_name = "estimate.log",
                                      n_loops = 3, 
                                      approximate_single_year_obs = T)
  saveRDS(weights, file = file.path(working_dir, reweight_folder, "Weights.RDS"))
}
## get reweighted MPDs to observe the effect
MPD_list = extract_reweighted_mpds(file.path(working_dir,reweight_folder))

## plot SSBs
plot_derived_quantities(MPD_list)
plot_fishery(MPD_list, quantity = "exploitation")
plot_recruitment(MPD_list, quantity = "ycs_values")
```

### Model quantities 

#### Fishing Pressures {-}
Below illustrates code to plot fishing pressure, but you can also easily adapt the code to plot catches. One thing to note, is Casal2 will report both `exploitation_rate` and `fishing_pressures`. For models that only have a single fishery per time-step or area these will be the same. If there are multiple fisheries interacting with the partition then they will differ. Fishing pressure is the maximum exploitation applied to the partition for that time-step and area. See the user manual for more detail on the difference. `exploitation_rate` reported is just
\[
\frac{catch}{vunerable}
\]

Some R-code used to summarise fishing pressures.

```{r pressures, fig.width=6, fig.height=4, results = 'hide', message=FALSE, warning=FALSE}
file_name = system.file("extdata", "SimpleTestModel", "estimate.log", 
                        package = "r4Casal2", mustWork = TRUE)
mpd = extract.mpd(file = file_name)
# Report labels
# names(mpd) 
# plot fishing pressures
fishery_info = get_fisheries(mpd)
head(fishery_info)
# Note this will print both fishing pressure and exploitation
my_plot = ggplot(fishery_info, aes(x = year, y = exploitation, col = fishery, linetype = fishery)) +
                   geom_line(size =2)
my_plot
```
Flexibility using standard ggplot functions

```{r pressures_alt, fig.width=6, fig.height=4}
# you can add adjust it as you please, for example if you want 3 panels for each fishery
my_plot + 
  facet_wrap(~fishery) + 
  theme(axis.text.x = element_text(angle = 90))
```

```{r pressures_alt_again, fig.width=6, fig.height=4}
# Adjust ylim and add a reference limit
my_plot + ylim(0,0.09) + geom_hline(yintercept = 0.05, linetype = "dashed")
```

#### Plotting selectivities
```{r selectivities_mpd, fig.width=6, fig.height=4}
selectivity_df = get_selectivities(model = mpd)
ggplot(selectivity_df, aes(x = bin, y = selectivity, col = label)) +
  geom_line(size = 1.5) +
  facet_wrap(~label)

```

## Multiple Casal2 runs with -i or -s 
```{r pressures_multi, fig.width=6, fig.height=4, results = 'hide', message=FALSE, warning=FALSE}
file_name = system.file("extdata", "SimpleTestModel","multi_run.log", 
                        package = "r4Casal2", mustWork = TRUE)
mpd = extract.mpd(file = file_name)
# Report labels
# names(mpd) 
# plot fishing pressures
fishery_df = get_fisheries(model = mpd)
my_plot = ggplot(fishery_df, aes(x = year, y = exploitation, col = factor(par_set))) +
  geom_line(size = 1.4) +
  facet_wrap(~fishery)
my_plot = my_plot + theme(axis.text.x = element_text(angle = 90))
# this will generate a generic ggplot
print(my_plot)
```

### Plotting selectivities {-}
```{r selectivities_multi, fig.width=8, fig.height=8}
selectivity_df = get_selectivities(mpd)
selectivity_df$par_set = factor(selectivity_df$par_set, ordered = T)
ggplot(selectivity_df, aes(x = bin, y = selectivity, col = report_label, line_type = par_set)) +
  geom_line(size = 1.5) +
  facet_grid(par_set~report_label)

```

### Plotting Fits {-}
```{r plot_relative_index, fig.width=8, fig.height=6}
my_plot = plot_relative_index(model = mpd, report_labels = c("chatTANbiomass"), plot.it = T, plot_type = "classic")
my_plot

```
